Analysing the Performance of Shallow-Water Models with the Roofline Model

### 2016-07-11 docs/papers/roofline_model/gocean_roofline.tex
roofline: rename roofline_haswell to roofline_desktop

In Figure~\ref{FIG_roofline_haswell} we plot the Roofline model
obtained for the E5-1620 CPU. In order to investigate the poor
performance of the u-Momentum kernel of NEMOLite2D, we also show the
performance of the (simpler) Continuity kernel from that code. As we
would expect, the performance of the u-Momentum kernel remains poor on
this CPU. For the $512^2$ domain the Continuity kernel performs well,
reaching the limit imposed by the bandwidth to main memory. However,
the working-set sizes for this kernel for the $128^2$ and $256^2$
domains are approximately 1.1~MB and 4.7~MB, respectively and
therefore both of these should fit comfortably within L3
cache. Despite this, the performance achieved with these domain sizes
falls short of the L3 bandwidth limit.  Even when we consider that the
Continuity kernel is just one of several within the time-stepping loop
of NEMOLite2D we see that the L3 bandwidth limit is still applicable;
the loop as a whole has a working-set size of $180\times n^{2}$
bytes. For the $32^2$ domain this is 180~KB ({\it i.e.}\ within L2),
while the $52^2$ (475~KB) through to the $128^2$ domain (2.9~MB) fall
well within the capacity of L3.

\begin{figure}
\centering
\includegraphics[width=120mm]{roofline_haswell}
\caption{Comparison of the performance achieved by different
  NEMOLite2D kernels on the E5-1620 CPU.}
\label{FIG_roofline_haswell}
\end{figure}

@@

The performance of kernel 5 is plotted using patterned bars in
Figure~\ref{FIG_daxpy_roofline}. This kernel has the same AI as kernel
4 but has an un-aligned array access. Comparing the performance of
these two kernels we see that the resulting reduction in the
efficiency with which the kernel can be SIMD vectorised (because of
the need for gather/scatter operations) harms performance,
particularly when the cost of accessing memory is small (i.e. when the
problem is contained in L1 or L2). Since all of the NEMOLite2D kernels
contain un-aligned array accesses, this goes some way to accounting
for the fact that they do not reach the theoretical limits of the
roofline plot in Figure~\ref{FIG_roofline_haswell}.

>>>>

In Figure~\ref{FIG_roofline_desktop} we plot the Roofline model
obtained for the E5-1620 CPU. In order to investigate the poor
performance of the u-Momentum kernel of NEMOLite2D, we also show the
performance of the (simpler) Continuity kernel from that code. As we
would expect, the performance of the u-Momentum kernel remains poor on
this CPU. For the $512^2$ domain the Continuity kernel performs well,
reaching the limit imposed by the bandwidth to main memory. However,
the working-set sizes for this kernel for the $128^2$ and $256^2$
domains are approximately 1.1~MB and 4.7~MB, respectively and
therefore both of these should fit comfortably within L3
cache. Despite this, the performance achieved with these domain sizes
falls short of the L3 bandwidth limit.  Even when we consider that the
Continuity kernel is just one of several within the time-stepping loop
of NEMOLite2D we see that the L3 bandwidth limit is still applicable;
the loop as a whole has a working-set size of $180\times n^{2}$
bytes. For the $32^2$ domain this is 180~KB ({\it i.e.}\ within L2),
while the $52^2$ (475~KB) through to the $128^2$ domain (2.9~MB) fall
well within the capacity of L3.

\begin{figure}
\centering
\includegraphics[width=120mm]{roofline_desktop}
\caption{Comparison of the performance achieved by different
  NEMOLite2D kernels on the E5-1620 CPU.}
\label{FIG_roofline_desktop}
\end{figure}

@@

The performance of kernel 5 is plotted using patterned bars in
Figure~\ref{FIG_daxpy_roofline}. This kernel has the same AI as kernel
4 but has an un-aligned array access. Comparing the performance of
these two kernels we see that the resulting reduction in the
efficiency with which the kernel can be SIMD vectorised (because of
the need for gather/scatter operations) harms performance,
particularly when the cost of accessing memory is small (i.e. when the
problem is contained in L1 or L2). Since all of the NEMOLite2D kernels
contain un-aligned array accesses, this goes some way to accounting
for the fact that they do not reach the theoretical limits of the
roofline plot in Figure~\ref{FIG_roofline_desktop}.
